Histo of ____

### 2017-07-13 msk2.rmd
updated readme and added sc

-  Gout 
-  Septic arthropathies
-  Trauma
-  Osteoarthritis
-  Psoriatic arthritis

>>>>

-  Gout: first MTP joint
-  Septic arthritis: one large joint
-  Trauma
-  Osteoarthritis: knee/hips (weight bearing)
-  RA: little joints
-  Psoriatic arthritis
-  Bleeding into the joints: haemophilia
-  Bite

Other Qs to ask

-  Skin conditions: psoriatic
-  Fever/inf: septic arthritis

Ex:

-  General inspection: how she looks, gait, how she is walking
-  Look: inflammed
-  Touch
-  Move
-  Maybe CVS and look at peripheral pulses

Pathophysiology of gout:

-  Hyperuricaemia: underexcretion (90%) or overproduction (10%)
-  Excess urate => crystalisation => inflamm => phago of crystals => inflamm: joints get affected
-  Can clear up in a couple of days

Risk factors

-  M
-  Obesity
-  Kidney disease

Diagnostic finding in pt with gout: urate crystals

Mx of gout:

-  Acute
    -  NSAIDs
    -  Prednisone
    -  Colchicine for acute gout: interrupts cycle of mono-Na crystal deposition. Stops the inflammatory reaction to the crystals. 
-  Chronic
    -  Lower urate to $<0.36\text{mmol/L}$
    -  Lifestyle modifications: low purine diet, low sugar (fructose => hyperuricaemia), stay hydrated
    -  Allopurinol: inhibits xanthine oxidase => inhibits urate production. Not really given during acute flare up because it *may* exacerbates.

Synovial fluid in following disorders:

-  Normal: clear straw like colour
-  Septic arthritis: elevated WCC, culture positive, purulent exudate
-  Crystal deposition disease: see crystals, elevated WCC, negative culture, high urate/Ca
-  RA: elevated WCC, sterile, normal urate, neutrophil predominance
-  OA: joint mice

Long term gout damage:

-  Joint damage
-  Recurrent gout: deposits
-  Skin nodules
-  Extra-articular damage: 
    -  Kidney injuries/stones
    -  HTN
    -  Hypercholesterolaemia/hyperuricaemia

Mx in this pt:

-  Check medications, maybe can't use the thiazide
-  Lifestyle mentioned above
